Unions of Balls and Configuration Space

### 2014-08-18 Writing/Final/article.tex
writing update

\title{ Determine Paths in the Same Homotopy Class }
\author{
        Yinan Zhang\\
        Department of Computer Science\\
        Dartmouth College\\
        Hanover, New Hampshire 03755, US
        \and
        Devin Balkcom \\
        Department of Computer Science\\
        Dartmouth College\\
        Hanover, New Hampshire 03755, US
}
\date{\today}

@@

\begin{abstract}
Describing the  \ldots
\end{abstract}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Introducing Our Algorithm
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph{A} \emph{The Algorithm outline} \hfill \\
\indent Before starting introducing our algorithm, we firstly clarify some preliminary tools we can have: We are sampling in $(X, \rho)$ metric space. Let $x$ be a configuration in $X$(usually \emph{C-space}), $\rho(x)$ gives the clearance of configuration $x$. By \emph{"clearance"}, we mean its (estimated) distance to the nearest obstacles. Ball $B(x, \rho(x))$ is referred as a $"sample"$ in the rest of the section. We make no assumption of the shape of obstacles regions in \emph{C-space}.

@@

    \indent \State $randomConfigs \gets SampleRandomConfigurations( \emph{C-space}, number )$

    \indent \State $maSamples \gets SampleOnMedialAxis( randomConfigs )$
    \indent \State $regSamples \gets SampleLeftArea(maSamples, randomConfigs)$

    \indent \State $dualShape \gets WeightedAlphaShape( maSmaples )$
    \indent \State $topologyRoadmap \gets Retract(dualShape)$

    \indent \State $brokenRoads \gets BreakGraph( topologyRoadmap )$

@@

\paragraph{B} \emph{Sampling in C-space} \hfill \\
\indent As pointed out in Figure \ref{fig:Algo}, this phase has two parts: Sampling on medial axis, and sampling the rest of free space.

@@

\paragraph{C} \emph{Constructing Topology Roadmap of C-space} \hfill \\

\paragraph{D} \emph{Breaking A Graph into Loop-free Parts} \hfill\\

\paragraph{E} \emph{Determine Paths Homotopy Class} \hfill \\

\section{Good Properties}\label{properties}

\section{Experiments}\label{experiments}
We now describe the some experiments and their results.

\section{Conclusions and Future Work}\label{conclusions}
We worked hard, and achieved very little.

@@

  \bibitem{Book} Steven M. LaValle, "Planning Algorithms", Cambridge University Press, ISBN 0-521-86205-1, 2006.
  \bibitem{UMAPRM} Yeh, Hsin-Yi Cindy, et al. "UMAPRM: Uniformly Sampling the Medial Axis."
  \bibitem{RRT} LaValle, Steven M. "Rapidly-Exploring Random Trees A New Tool for Path Planning." (1998).
  \bibitem{PRM} L. E. Kavraki, P. Svestka, L. E. K. P. Vestka, J. claude Latombe, and M. H. Overmars, “Probabilistic roadmaps for path planning in high-dimensional configuration spaces,” IEEE Trans. Robot. Autom., vol. 12, pp. 566–580, 1996.
  \bibitem{MAPRM} Wilmarth, Steven A., Nancy M. Amato, and Peter F. Stiller. "MAPRM: A probabilistic roadmap planner with sampling on the medial axis of the free space." Robotics and Automation, 1999. Proceedings. 1999 IEEE International Conference on. Vol. 2. IEEE, 1999.
  \bibitem{DualShape} Edelsbrunner, Herbert. "The union of balls and its dual shape." Proceedings of the ninth annual symposium on Computational geometry. ACM, 1993.
  \bibitem{PointSetShape} Edelsbrunner, Herbert; Kirkpatrick, David G.; Seidel, Raimund (1983), "On the shape of a set of points in the plane", IEEE Transactions on Information Theory 29 (4): 551–559, doi:10.1109/TIT.1983.1056714.
  \bibitem{PathNonexistance} Z. McCarthy. Bretl, and S. Hutchinson, "Proving path non-existence using sampling and alpha shapes," in IEEE International Conference on Robotics and Automation (ICRA), May 2012.
  \bibitem{betti wiki} "Betti Number", From Wikipedia -- The Free Encyclopedia. ( $http://en.wikipedia.org/wiki/Betti_number$ )    
  \bibitem{Edel Book}Edelsbrunner, Herbert, and John Harer. "Computational topology: an introduction." American Mathematical Soc., 2010. page. 88 - 93.
  \bibitem{alpha betti} Robins, Vanessa. "Computational topology for point data: Betti numbers of α-shapes." Morphology of Condensed Matter. Springer Berlin Heidelberg, 2002. 261-274.
\end{thebibliography}

>>>>

\title{ Unions of Balls and Configuration Space }
\author{
        Yinan Zhang\\
        Department of Computer Science\\
        Dartmouth College\\
        Hanover, New Hampshire 03755, US
        \and
        Devin Balkcom \\
        Department of Computer Science\\
        Dartmouth College\\
        Hanover, New Hampshire 03755, US
}
\date{\today}

@@

\begin{abstract}
Covering both the configuration space and represent the right topology is hard to achieve at the same time with traditional sampling-based motion planning algorithms, such as Probabilistic Roadmap and Rapidly-Exploring Random Tree. In this paper, we proposed a method that is able to cover near optimal path and study the shape and topology of configuration space. 
\end{abstract}

@@

\indent The first betti number of a graph is very easy to compute. Given a graph $G$, with $n$ vertices, $m$ edges and $k$ components. It's 1st betti number is:

$\beta_1 = m - n + k$

\indent This can be proved straightforwardly by mathematical induction on the number of edges. A new edge either increments the number of 1-cycles or decrements the number of connected components. \cite{betti wiki}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Introduction to Our Algorithm
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%====================================
% Algorithm outline
%====================================
\paragraph{A} \emph{The Algorithm outline} \hfill \\
\indent Before starting introducing our algorithm, we firstly clarify some preliminary tools we can have: We are sampling in $(X, \rho)$ metric space. Let $x$ be a configuration in $X$(usually \emph{C-space}), $\rho(x)$ gives the clearance of configuration $x$. By \emph{"clearance"}, we mean its (estimated) distance to the nearest obstacles. Ball $B(x, \rho(x))$ is referred as a $"sample"$ in the rest of the section. We make no assumption of the shape of obstacles regions in \emph{C-space}.

@@

    // Phase One
    \indent \State $randomConfigs \gets SampleRandomConfigurations( \emph{C-space}, number )$

    // Phase Two
    \indent \State $maSamples \gets SampleOnMedialAxis( randomConfigs )$
    \indent \State $regSamples \gets SampleLeftArea(maSamples, randomConfigs)$

    // Phase Three
    \indent \State $dualShape \gets WeightedAlphaShape( some smaples )$
    \indent \State $topologyRoadmap \gets Retract(dualShape)$

    // Phase Four
    \indent \State $brokenRoads \gets BreakGraph( topologyRoadmap )$

@@

%====================================
% Covering c-space
%====================================
\paragraph{B} \emph{Sampling in C-space} \hfill \\
\indent As pointed out in Figure \ref{fig:Algo}, this phase has two parts: Sampling on medial axis, and sampling the rest of free space.

@@

\indent The reason we sample on medial axis first is that 1. these balls are maximal ones as the definition of medial axis suggests, and 2. medial axis $MA(C_{free})$ is a SDR of \emph{C-space}. \cite{MAPRM2}

\indent Figure \ref{SampleLeftArea} describes how we sample area not yet covered by balls centered on medial axis. The basic idea behind the algorithm is to keep sampling new balls on the boundary of existing ones by "pushing" random configurations inside existing balls to the boundary of them and determine if they are in any other balls. We restrict that all balls to be larger than a minimal radius ball, thus void getting a lot of very small samples.

\indent Getting configurations inside a ball doesn't have to be $O(n)$ costly. It can be done using KNN algorithm, which is more efficient than iterating everything. Since we generated random configuration $n$ times, not all are feasible, in a 2D space with area $A$. We expect to have $n/A$ configurations in an unit area. A ball with radius $r$ has area $\pi r^2$, approximately $k = \frac{n \pi r^2}{A}$ configurations are inside the ball. We then project these (uniformly distributed) configurations to the boundary of the ball. Actually, choosing a larger $k$ could be better: we virtually extend the radius of our discs, which results in a smaller density in arcs that are very close to obstacles than those far from obstacles. Samples got in this part is called \emph{regular samples}.

\indent Figure \ref{Samples} is a result of this phase running in a rectangle world with 3 obstacles in the middle. Gray balls are centered on medial axis, while green ones are not. Red points are centers.

\begin{figure}
\begin{algorithmic}[1] 
  \Function{SampleLeftArea}{ maSamples, randomConfigs }
  	\State $regSamples \gets EmptySet$ \Comment{Set of balls not on MA}
  	\State $newBall \gets$ True	\Comment{If we can get new balls}
  	\State $lastRoundBalls \gets maSamples$
  	\While{$newBall$}
  		\State $newBall \gets False$
		\State $bndConfigs \gets EmptySet$ \Comment{A set of ball boundary configs}  		
  		\For {$ball \in lastRoundBalls$}
  			\State $innerConfigs \gets KNearest(ball, randomConfigs)$; \Comment{KNN}
  			\State $bndConfigs.append( getBoundaryConfigs(ball, innerConfigs) )$
  		\EndFor
  		\State $lastRoundBalls \gets EmptySet$
  		\For{$config \in bndConfigs$}
  			\If{$config$ not inside any existing ball} \Comment{ $O(n^2)$ time}
  				\State $ball \gets getNewBall(config)$ \Comment{get a new ball}
  				\If{$ball.radius \geq min\_radius$}
  					\State $newBall \gets True$
  					\State $lastRoundBalls.append( ball )$
  					\State $regSamples.append(ball);$
  				\EndIf
  			\EndIf
  		\EndFor
  	\EndWhile
  	\State \Return $regSamples$
  \EndFunction
  \end{algorithmic}
  \caption{\label{SampleLeftArea} Algorithm to sample left free area}
\end{figure}

\begin{figure}
	\center{\includegraphics[scale=0.3]{./pics/Samples.PNG}}
	\caption{\label{Samples} Samples got by the algorithm}
\end{figure}

%====================================
% Dual shape and topology roadmap
%====================================

\paragraph{C} \emph{Constructing Topology Roadmap of C-space} \hfill \\
\indent After getting these samples, we can now start to build a structure that captures the topology of the space.

\indent This phase, like the previous one, has two parts: 1. Construct the dual shape of medial axis samples got from phase one. 2. Simplify the dual shape to a simple graph.

\indent In the first part, constructing the dual shape using only the medial axis sample could break the topology of medial axis. The reason is shown in Figure \ref{badthing}: three balls are all centered on medial axis, but their dual shape will be an empty triangle, which introduces a "hole" in that area. To solve this problem we can try to sample a ball centered at the the radical center after getting an empty triangle, and rebuild the dual shape for these four balls. \cite{radical center} Proving the solution is very simple: if there is no obstacles in that area, the newly sampled ball will cover the space, if there is obstacle, the ball will either not be sampled or not connect three of the balls.

\indent With these balls, we then compute their dual shape, which is formed by some 1-simplices and 2-simplices in 2D. The runtime of computing dual shape will be $O(n^2)$ which is the worst case when doing regular triangulation of weighted points. Although using all balls will avoid problem as in Figure \ref{badthing}, we are making $n$ much larger, which results in a bad runtime of this step.

\indent One might argue that we can simply get configurations on medial and connect them as a roadmap. However, this roadmap doesn't maintain the topology of \emph{C-space}. At some corners, we will still get triangles which don't have information if there are obstacles inside the triangle. By building the dual shape, this will no longer be a problem.

\indent The second part of this phase can be stated as: retract all connected 2-simplices into one point, reconnect these retracted points with other ones using their previous connection.

\begin{figure}
	\center{\includegraphics[scale=1]{./pics/badthing.pdf}}
	\caption{\label{badthing}} An example that medial axis samples break the topology structure of the medial axis.
\end{figure}

\begin{figure}
	\center{\includegraphics[scale=0.3]{./pics/my_dualshape.PNG}}
	\caption{\label{ourdualshape}} Dual shape of balls generated in phase 1 (centered on medial axis).
\end{figure}

\begin{figure}
	\center{\includegraphics[scale=0.4]{./pics/simplified_graph.png}}
	\caption{\label{simplifiedgraph}} Simplified graph of the dual shape in Figure \ref{ourdualshape}.
\end{figure}

%====================================
% Break graph
%====================================
\paragraph{D} \emph{Breaking A Graph into Loop-free Parts} \hfill\\
\indent After getting the topology roadmap, we technically already have a planner, but still we can simplify it. We now discuss breaking a graph into several components, such that no two components can merge into a new graph without forming a loop. By doing this, a graph is even simplified. Knowing which components are the start and goal configurations in, we can find a path connecting these components immediately. This technique can be a great improve when $C-space$ is very large and traditional \emph{PRM} or \emph{RRT} methods will take a long time to solve.

\indent Let $g_1$ and $g_2$ be two graphs without loops. We say $g_1$ and $g_2$ to be mergeable if: 1. they share at least one common node, 2. the new graph formed by these two graph has no loop.

\indent Determine if a simple graph has a loop is very easy using the first Betti number $\beta_1$ mentioned in section 2.

\indent Figure \ref{breakgraph} shows our algorithm to break a graph. The idea behind this algorithm is to take every edges as an initial component, then iterate each component to find and merge with a mergeable component until there is no two mergeable components.

\begin{figure}
\begin{algorithmic}[1] 
  \Function{BreakGraph}{ graph }
  	\State $components \gets EmptySet$
  	\For {$edge \in graph.edges$}
  		\State components.append( new Graph(edge) )
  	\EndFor

  	\State \Return MergeComponents( components )
  \EndFunction
\end{algorithmic}

\begin{algorithmic}[1]   
  \Function{MergeComponents}{components}
  	\State $pq \gets PriorityQueue()$
    \For{$component \in components$}
        \State $pq.push(component, component.size())$
    \EndFor
    \While{ $pq.size() > 1$ }
    	\State $smallest\_component \gets pq.pop()$
    	\State $mergeable\_component \gets$ \emph{find first mergeable component in pq}
		\State $new\_component \gets smallest\_component.merge\_with( mergeable\_component )$
    	\State $components.remove( smallest\_component )$
    	\State $components.remove( mergeable\_component )$
    	\State $pq.remove( mergeable\_component )$
    	\State $components.append( new\_component )$
    	\State $pq.push(new\_component, new\_component.size())$
	\EndWhile

    \State \Return $components$
  \EndFunction  
\end{algorithmic}
	\caption{\label{breakgraph}} Breaking a graph into several loop-free components.
\end{figure}

\begin{figure}
	{\includegraphics[scale=0.7]{./pics/initial.png}}
	{\includegraphics[scale=0.7]{./pics/broken.png}}

	{\includegraphics[scale=0.65]{./pics/initial2.png}}
	{\includegraphics[scale=0.65]{./pics/broken2.png}}\
	\caption{\label{brokengraphs} Some examples of breaking a graph. Graphs are visualized using $d3.js$ $(http://www.d3js.org)$ }
\end{figure}

%====================================
% Path topology
%====================================
\paragraph{E} \emph{Determine Paths Homotopy Class} \hfill \\
\indent A continuous function, or a map, $f: X \rightarrow Y$ is called a \emph{homotopy equivalence} if there is a map $g: X \rightarrow Y$ such that there exist a continuous function $H: X \times [0,1] \rightarrow Y$ with $H(x, 0) = f$ and $H(x,1) = g$.

\indent Besides using those components to plan motions for a large \emph{C-space}, they can also be used to determine if two paths are in the same homotopy class.

\indent Let $\sigma_1$ and $\sigma_2$ be two paths, with the same start and goal configuration, going through from components $g_1$ to $g_2$ via two nodes $A$ and $B$. $\sigma_1$ and $\sigma_2$ are homotopy equivalent if $A = B$.

\indent Proof can be found in section 4.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Properties of Our Algorithm
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Good Properties}\label{properties}

  \begin{theorem}
  Let $S_{init}$ be the set of initial random configurations. With minimum radius of $r_{min}$, configurations with clearance larger than $r_{min}$ will be covered, when $|S_{init}| \rightarrow \infty$.
  \end{theorem}

  \paragraph{\emph{Proof:}} 
  \indent When the number of initial random configurations approaches infinite, the boundaries of all existing balls are fully covered. So, every configuration not inside any existing balls will be samples as a ball, unless it's less than $r_{min}$ clearance.

  \begin{theorem}
  Let $B$ be the set of ball samples got in phase one. Then $|B|$ is finite.   
  \end{theorem}
  \paragraph{\emph{Proof:}} 
  \indent We restrict balls to have no less than $r_{min}$ radius and new balls to be centered outside existing ones. Assume we partition the space into $r_{min} \cdot r_{min}$ cells in 2D. The number of cells is finite, if the space is finite. Putting balls with exactly $r_{min}$ radius in the centers of these cells will cover the space totally. The number of balls is finite. Phase one will generate balls no more than the number of balls generated in this way. Figure \ref{ourdualshape} shows the dual shape of balls generated by previous phase.

  \begin{theorem}
  Let $b \in B$ be an element of $B$, $A$ and $B$ are two configurations in \emph{C-space}. The optimal path $\sigma$ between $A$ and $B$ is in $\bigcup_{b\in B}$ if $\sigma$ has clearance no less than $r_min$. 
  \end{theorem}   
  \paragraph{\emph{Proof}}
  \indent This is a direct conclusion of previous theorem.

  \begin{theorem}
  Any path $\sigma$ can be continuously deformed into medial axis. 
  \end{theorem}   
  \paragraph{\emph{Proof}}
  \indent This is the definition of Medial Axis $MA(C_{free})$ being a strong deformation retraction of $C_{free}$.

  \begin{theorem}
  Given path $\sigma_1$ and $\sigma_2$, we say they have the same homotopy type if their image in the topology roadmap is the same. 
  \end{theorem}
  \paragraph{\emph{Proof}}
  \indent Let $x$ be a point in the space $X$, $B = \{ b_1, b_2, ..., b_n \}$ be the set of balls centered on medial axis. Define map $H(x)$:

  $H(x) = b_i \indent, if \indent |x-b| < |x-b'|, b' \in B$

  \indent This continuous function maps any point in the space to balls centered on medial axis. Dual shape maps any points in the balls to the topology roadmap we build.

  \indent Therefore, if two paths have the same image in the topology roadmap. They have the same homotopy type.

  \begin{theorem}
  Let $g_1$ and $g_2$ be two loop-free graphs such that merging their common nodes will produce a new graph with loops, $\sigma_1$ and $\sigma_2$, with the same start and goal nodes, be two paths going from $g_1$ to $g_2$ via common nodes $A$ and $B$. $\sigma_1$ and $\sigma_2$ have different homotopy type if $A \neq B$
  \end{theorem}
  \paragraph{\emph{Proof}}
  \indent Because $A$ and $B$ are two common nodes shared by $g_1$ and $g_2$, if $A \neq B$, merging $g_1$ and $g_2$ will create a graph with a loop going from $A \rightarrow B \rightarrow A$. Since $\sigma_1$ and $\sigma_2$ have the same start and goal nodes, merging them will also create a loop. So $\sigma_1$ and $\sigma_2$ have different homotopy type.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Experiments
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Experiments}\label{experiments}
%We now describe the some experiments and their results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	Conclusions and Future work
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}\label{conclusions}
We worked hard, and achieved very little.

@@

  \bibitem{Book} Steven M. LaValle, "Planning Algorithms", Cambridge University Press, ISBN 0-521-86205-1, 2006.
  \bibitem{UMAPRM} Yeh, Hsin-Yi Cindy, et al. "UMAPRM: Uniformly Sampling the Medial Axis."
  \bibitem{RRT} LaValle, Steven M. "Rapidly-Exploring Random Trees A New Tool for Path Planning." (1998).
  \bibitem{PRM} L. E. Kavraki, P. Svestka, L. E. K. P. Vestka, J. claude Latombe, and M. H. Overmars, “Probabilistic roadmaps for path planning in high-dimensional configuration spaces,” IEEE Trans. Robot. Autom., vol. 12, pp. 566–580, 1996.
  \bibitem{MAPRM} Wilmarth, Steven A., Nancy M. Amato, and Peter F. Stiller. "MAPRM: A probabilistic roadmap planner with sampling on the medial axis of the free space." Robotics and Automation, 1999. Proceedings. 1999 IEEE International Conference on. Vol. 2. IEEE, 1999.
  \bibitem{MAPRM2} S. A. Wilmarth, N. M. Amato, and P. F. Stiller. MAPRM: A probabilistic roadmap planner with sampling on the medial axis of the free space. Technical Report TR98-022, Department of Computer Science, Texas A\&M University, College Station, TX, Nov. 1998.
  \bibitem{DualShape} Edelsbrunner, Herbert. "The union of balls and its dual shape." Proceedings of the ninth annual symposium on Computational geometry. ACM, 1993.
  \bibitem{PointSetShape} Edelsbrunner, Herbert; Kirkpatrick, David G.; Seidel, Raimund (1983), "On the shape of a set of points in the plane", IEEE Transactions on Information Theory 29 (4): 551–559, doi:10.1109/TIT.1983.1056714.
  \bibitem{radical center} "Radical Center", From Wolfram MathWorld -- the web's most extensive mathematics resource. ($http://mathworld.wolfram.com/RadicalCenter.html$)
  \bibitem{PathNonexistance} Z. McCarthy. Bretl, and S. Hutchinson, "Proving path non-existence using sampling and alpha shapes," in IEEE International Conference on Robotics and Automation (ICRA), May 2012.
  \bibitem{betti wiki} "Betti Number", From Wikipedia -- The Free Encyclopedia. ( $http://en.wikipedia.org/wiki/Betti_number$ )    
  \bibitem{Edel Book}Edelsbrunner, Herbert, and John Harer. "Computational topology: an introduction." American Mathematical Soc., 2010. page. 88 - 93.
  \bibitem{alpha betti} Robins, Vanessa. "Computational topology for point data: Betti numbers of α-shapes." Morphology of Condensed Matter. Springer Berlin Heidelberg, 2002. 261-274.
\end{thebibliography}
